A weakly structured stem for human origins in Africa

### 2023-03-04 paper.tex
Update paper.tex

Third, multiple studies have shown a correspondence between phenotypic
differentiation, usually assessed with measurements of the cranium, and genetic
differentiation among human populations and between humans and Neanderthals
\citep{Relethford1994-mh,Weaver2008-ho,Von_Cramon-Taubadel2009-zb} (see also SI
Section~\ref{sec:morpho_diff}). This correspondence potentially allows
predictions of our model to be related to the fossil record. The fossil record
of Africa is sparse during the time period of the stems, but of the fossils
that date to this time period, some are fairly similar overall in morphology to
contemporary humans (e.g., Omo 1 from Omo Kibish, Ethiopia
\citep{Day1969-rh,Vidal2022-qe}), others are similar in some morphological
features to contemporary humans (e.g., Irhoud 1 from Jebel Irhoud, Morocco
\citep{Hublin2017-cq,Richter2017-zu}), and others are different enough in
morphology to have been assigned to species other than \emph{Homo sapiens}
(e.g., DH1 from Dinaledi, South Africa \citep{Berger2015-bq,Dirks2017-uk}). If,
as our model predicts, the genetic differences between the stems were
comparable to those among contemporary human populations, the most
morphologically divergent fossils are unlikely to represent branches that
contributed appreciably to contemporary human ancestries.

>>>>

Third, multiple studies have shown a correspondence between phenotypic differentiation, 
usually assessed with measurements of the cranium, and genetic differentiation among 
human populations and between humans and Neanderthals \citep{Relethford1994-mh,
Weaver2008-ho,Von_Cramon-Taubadel2009-zb} (see also SI Section~\ref{sec:morpho_diff}). 
This correspondence potentially allows predictions of our model to be related to the 
fossil record. \emph{Homo sapiens} fossils from Iho Eleru, Nigeria (13ka) 
\citep{Bergstrom2021-iw} \textcolor{red}{ADD REFERENCE} Harvati et al. 2011 ``The Later 
Stone Age Calvaria from Iwo Eleru, Nigeria: Morphology and Chronology'', Ishango, 
Democratic Republic of Congo (25-20ka) \textcolor{red}{ADD REFERENCE} Crevecoeur et al. 
2016 ``Late Stone Age Human Remains from Ishango (Democratic Republic of Congo): New 
Insights on Late Pleistocene Modern Human Diversity in Africa'', and Nazlet Khater, 
Egypt (40-35ka) \textcolor{red}{ADD REFERENCE} Crevecoeur 2012 ``The Upper 
Paleolithic Human Remains of Nazlet Khater 2 (Egypt) and Past Modern Human Diversity'' 
have morphological features that may indicate gene flow from archaic hominins. If 
this interpretation is correct, these fossils are unlikely to represent populations 
ancestral to contemporary human populations, because we would have expected to detect 
admixture from late surviving archaic hominins. In our multiple-merger model, we do 
infer late gene flow from one of the stems into West African populations (25ka), but 
this is not admixture from an archaic hominin lineage. The fossil record of Africa is 
sparse during the time period of the stems, but of the fossils that date to this time 
period, some are fairly similar overall in morphology to contemporary humans (e.g., 
Omo 1 from Omo Kibish, Ethiopia \citep{Day1969-rh,Vidal2022-qe}), others are similar 
in some morphological features to contemporary humans (e.g., Irhoud 1 from Jebel Irhoud, 
Morocco \citep{Hublin2017-cq,Richter2017-zu}), and others are different enough in 
morphology to have been assigned to species other than \emph{Homo sapiens} (e.g., DH1 
from Dinaledi, South Africa \citep{Berger2015-bq,Dirks2017-uk}). If, as our model 
predicts, the genetic differences between the stems were comparable to those among 
contemporary human populations, the most morphologically divergent fossils are unlikely 
to represent branches that contributed appreciably to contemporary human ancestries.
